---
title: "12/10/2020 Example"
author: "Kevin Donovan"
date: "12/10/2020"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r packages}
library(tidyverse)
library(readr)
library(gtsummary)
library(gt)
library(flextable)
library(rmarkdown)
library(DT)
```

# Introduction

This an example of using R Markdown with IBIS data.

```{r data_management}
ibis_data <- read_csv("../Data/Cross-sec_full.csv", na = c(".","","NA"))
```

# Summary Statistics
```{r summ_stats}
tf <- tbl_summary(ibis_data, by="GROUP") %>%
  as_flex_table()

tf

save_as_docx(tf, path="summ_stats.docx")
```

# Interactive Tables
```{r int_table}
paged_table(ibis_data)

datatable(ibis_data)
```

# First Analysis
First we visualize AOSI total score and Mullen composite score at 6 months using a 2D scatterplot.  We also color by diagnosis, panel by gender, and we include trend lines for the whole sample and by gender.

```{r v06_aosi_mullen_plot}
plot(x=ibis_data$`V06 aosi,total_score_1_18`, 
     y=ibis_data$`V06 mullen,composite_standard_score`)

ggplot(data=ibis_data, mapping=aes(x=`V06 aosi,total_score_1_18`,
                                   y=`V06 mullen,composite_standard_score`,
                                   color=SSM_ASD_v24))+
  geom_point()+
  geom_smooth(se=FALSE)+
  geom_smooth(color="blue")+
  facet_grid(~Gender)+
  labs(title="My title", subtitle = "My subtitle", caption="My footer",
       color=" 24 month ASD diagnosis")+
  xlab("6 month AOSI\nTotal Score")+
  ylab("6 month MSEL\nComposite Std. Score")
```

# Additional Tables Resources
Links for help with these table packages:

[Basic help](https://rstudio.github.io/distill/tables.html)

[datatable](https://rstudio.github.io/DT/)

[gt](https://gt.rstudio.com/)

[flextable](https://davidgohel.github.io/flextable/index.html)

[gtsummary](http://www.danieldsjoberg.com/gtsummary/)