---
title: "Presenting Regression Analyses in R"
author: "Kevin Donovan"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---
```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, message = FALSE, warning = FALSE, fig.width = 8, 
                      fig.height = 4)

library(tidyverse)
library(shiny)
library(rmarkdown)
library(broom)
library(gtsummary)
library(flextable)
library(ggpubr)
library(nlme)
library(lme4)
library(broom.mixed)
library(GGally)
library(corrplot)
library(ggcorrplot)
library(lsmeans)
library(sjPlot)
```

```{r}
brain_data <- read_csv("../Data/IBIS_brain_data_ex.csv") %>%
  mutate(V24_ASD_DX_edit = factor(ifelse(grepl("YES", V24_ASD_DX), "Positive",
                                  ifelse(grepl("NO", V24_ASD_DX), "Negative", NA))))

behav_data <- read_csv("../Data/Cross-sec_full.csv", na=c(".", "", " "))
names(behav_data) <- gsub(" |,",".",names(behav_data))

mixed_model_data <- behav_data %>%
  select(Identifiers, GROUP, Gender,
         V36.mullen.composite_standard_score:V12.mullen.composite_standard_score)

vars_to_convert <- names(mixed_model_data)[c(-1,-2,-3)]

mixed_model_data <- mixed_model_data %>%
  gather(variable, var_value, vars_to_convert) %>%
  separate(variable,c("Visit","Variable"),sep=3) %>% 
  spread(key=Variable, value=var_value) %>%
  plyr::rename(c(".mullen.composite_standard_score"="Mullen_Composite_Score")) %>%
  mutate(ASD_Diag = factor(ifelse(grepl("ASD", GROUP), "ASD_Neg", "ASD_Pos")),
         Visit=factor(Visit),
         Visit_Num=as.numeric(Visit)-1,
         Mullen_Composite_Score=as.numeric(Mullen_Composite_Score)) %>%
  arrange(Identifiers, Visit)
```

# Correlation analyses

Correlation heat map

```{r fig.width = 7, fig.height = 6}
brain_data_v24 <-
  brain_data %>%
  select(names(brain_data)[grepl("V24", names(brain_data))]) %>%
  select(EACSF_V24:RightAmygdala_V24 )

p.mat <- cor_pmat(brain_data_v24)

ggcorrplot(cor(x=brain_data_v24, method="pearson", use="pairwise.complete.obs"), 
           hc.order = TRUE, type = "lower", lab = TRUE, outline.col = "white",
           p.mat = p.mat)

corrplot(cor(x=brain_data_v24, method="pearson", use="pairwise.complete.obs"))
```

Grid of heat maps

```{r fig.width = 7, fig.height = 6}
brain_data_v12 <-
  brain_data %>%
  select(names(brain_data)[grepl("V12", names(brain_data))]) %>%
  select(EACSF_V12:RightAmygdala_V12 )

brain_data_v24 <-
  brain_data %>%
  select(names(brain_data)[grepl("V24", names(brain_data))]) %>%
  select(EACSF_V24:RightAmygdala_V24 )

p.mat_v12 <- cor_pmat(brain_data_v12)
p.mat_v24 <- cor_pmat(brain_data_v24)

heatmaps <- list(ggcorrplot(cor(x=p.mat_v12, method="pearson", use="pairwise.complete.obs"), 
           hc.order = TRUE, type = "lower", lab = TRUE, outline.col = "white",
           p.mat = p.mat)
           ,
           ggcorrplot(cor(x=p.mat_v24, method="pearson", use="pairwise.complete.obs"), 
           hc.order = TRUE, type = "lower", lab = TRUE, outline.col = "white",
           p.mat = p.mat))

ggarrange(plotlist=heatmaps, labels = c("12 months", "24 months"))
```

Lines of best fit with correlations

```{r fig.width = 7, fig.height = 6}
ggplot(mixed_model_data, mapping=aes(x=Visit_Num, y=Mullen_Composite_Score,
                                     color=ASD_Diag))+
  geom_point()+
  geom_smooth(method="lm", se=FALSE)+
  facet_grid(~Gender)+
  stat_cor()+
  theme_bw()
```

# Correlation analyses

Matrix of distribution summary visuals

```{r fig.width = 10, fig.height = 8}
brain_data_v24 <-
  brain_data %>%
  select(c(names(brain_data)[grepl("V24", names(brain_data))], "RiskGroup")) %>%
  select(EACSF_V24:RightAmygdala_V24, RiskGroup)

ggpairs(brain_data_v24, 
        columns = names(brain_data_v24)[!grepl("RiskGroup", names(brain_data_v24))],
        ggplot2::aes(colour=RiskGroup, alpha=0.25)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

# Summary statistics 
- Can create easily formatted summary stats tables **in code**

```{r fig.width = 10, fig.height = 8}
tbl_summary(data=brain_data_v24, by=RiskGroup,
            missing_text = "Missing",
            statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
  add_p(list(all_continuous() ~ "aov", 
             all_categorical() ~ "chisq")) %>%
  add_n() %>%
  as_flex_table() %>%
  bold(bold = TRUE, part = "header") %>%
  autofit()
```

# Regression summaries
## Tables
### Linear Models

```{r}
lm_fit <- lm(V24_MSEL_EL_ae~RiskGroup+RightAmygdala_V24+
               RiskGroup*RightAmygdala_V24, 
             data=brain_data)

tbl_regression(lm_fit,
               pvalue_fun = ~style_pvalue(.x, digits = 3),
               estimate_fun = ~style_number(.x, digits = 3)) %>%
  as_flex_table() %>%
  autofit()

# Can also do by hand using tidy()
lm_fit_tidy <- tidy(lm_fit)
lm_fit_tidy
```

### Generalized Linear Models

```{r}
glm_fit <- glm(V24_ASD_DX_edit~V24_MSEL_EL_ae+RightAmygdala_V24+Cand_Sex, 
               family=binomial(),
             data=brain_data)

tbl_regression(glm_fit,
               pvalue_fun = ~style_pvalue(.x, digits = 3),
               estimate_fun = ~style_number(.x, digits = 4)) %>%
  as_flex_table() %>%
  autofit()

tbl_regression(glm_fit,
               pvalue_fun = ~style_pvalue(.x, digits = 3),
               estimate_fun = ~style_number(.x, digits = 4),
               exponentiate = TRUE) %>%
  as_flex_table() %>%
  autofit()
```

### Mixed Models

```{r}
mixed_fit <-
  lme(Mullen_Composite_Score~Visit_Num + ASD_Diag, 
    data=mixed_model_data,
    random = ~1|Identifiers,
    na.action = na.omit)

# Can tidy using broom.mixed package
tidy(mixed_fit) %>%
  filter(effect=="fixed") %>%
  select(-effect, -group) %>%
  flextable()
```

## Visualizations
### Linear Models

```{r}
lm_fit <- lm(V24_MSEL_EL_ae~RiskGroup+RightAmygdala_V24+
               RiskGroup*RightAmygdala_V24, 
             data=brain_data)

model_data <- model.frame(lm_fit)
model_data$predict_MSEL <- predict(lm_fit, newdata=model_data)

# Now plot trend line
# Method 1
ggplot(data=model_data,
       mapping=aes(x=RightAmygdala_V24, y=V24_MSEL_EL_ae, color=RiskGroup))+
  geom_point()+
  geom_line(mapping=aes(y=predict_MSEL), size=1.5)+
  theme_bw()

# Method 2
ggplot(data=brain_data,
       mapping=aes(x=RightAmygdala_V24, y=V24_MSEL_EL_ae, color=RiskGroup))+
  geom_point()+
  geom_smooth(method="lm", se=FALSE, size=1.5)+
  theme_bw()

# Least square means
ls_means_df <- lsmeans(lm_fit, "RiskGroup",
                       at = list(RightAmygdala_V24=mean(brain_data$RightAmygdala_V24,
                                                        na.rm=TRUE))) %>% data.frame()

ggplot(data=ls_means_df,
       mapping=aes(x=RiskGroup, y=lsmean, color=RiskGroup))+
  geom_point()+
  geom_errorbar(mapping=aes(ymin=`lower.CL`, ymax=`upper.CL`))+
  theme_bw()
```

### Generalized Linear Models

```{r}
glm_fit <- glm(V24_ASD_DX_edit~V24_MSEL_EL_ae+RightAmygdala_V24+Cand_Sex, 
               family=binomial(),
             data=brain_data)

# Creating a forest plot
plot_model(glm_fit)

# Creating a forest plot by hand
glm_fit_tidy <- tidy(glm_fit)
confints_glm <- confint(glm_fit)

glm_fit_tidy <- cbind(glm_fit_tidy, confints_glm) %>%
  mutate(or_est = exp(estimate),
         or_lci = exp(`2.5 %`),
          or_uci = exp(`97.5 %`)) %>%
  filter(term!="(Intercept)")
```